Simulating High-Dimensional Multivariate Data

### 2021-05-25 020-simulation-algorithm.Rmd
working template
Revise for JSS submission up to Chapter 4

This section describes our methods involved in simulating a random vector $\bf Y$ with $Y_i$ components for $i=1,\ldots,d$. Each $Y_i$ has a specified marginal CDF $F_i$ and its inverse $F^{-1}_i$. To characterize dependency, every pair $(Y_i, Y_j)$ has a given Pearson correlation $\rho_P$, Spearman correlation $\rho_S$, and/or Kendall's $\tau$. The method can be described as a *high-performance Gaussian copula* (Equation \@ref(eq:gauss)) providing a high-dimensional NORTA-inspired algorithm.

@@

1. Simulate a random vector $\bf Z$ with $d$ *independent and identically distributed* (iid) standard normal components.
2. Determine the input matrix $\Sigma_{\bf Z}$ that corresponds with the specified output $\Sigma_{\bf Y}$.
3. Produce a Cholesky factor $M$ of $\Sigma_{\bf Z}$ such that $M M^{\prime}=\Sigma_{\bf Z}$.
4. Set $X$ by $X \gets MZ$.
5. $\text{Return} \; Y \; \text{where} \; Y_i \gets F_i^{-1}[\Phi(X_i)], \; i=1,...,d$.

@@

With modern parallel computing, steps 1, 3, 4, 5 are readily implemented as high-performance, multi-core and/or graphical-processing-unit (GPU) accelerated algorithms --- providing fast scalability using readily-available hardware.

Matching specified Pearson correlation coefficients exactly (step 2 above), however, is computationally costly. In general, there is no closed-form correspondence between the components of the input $\Sigma_{\bf Z}$ and target $\Sigma_{\bf Y}$. Matching the correlations involves evaluating or approximating $\binom{d}{2}$ integrals of the form

\begin{equation}
    \mathrm{E}\left[Y_i Y_j\right] = \int_{-\infty}^{\infty} \int_{-\infty}^{\infty} F_i^{-1}\left[\Phi(z_i)\right] F_j^{-1}\left[\Phi(z_j)\right] \phi(z_i, z_j, \rho_z) dz_i dz_j,
    (\#eq:pearsonIntegralRelation)
\end{equation}

where $\phi(\cdot)$ is the joint probability distribution function of two correlated standard normal variables. For HD data, (nearly) exact evaluation may become too costly to enable practical simulation studies. For low-dimensional problems, however, methods and tools exist to match Pearson correlations precisely; see [@Xia17] and the publicly available `nortaRA` R package [@Chen2001]. As described later, to enable HD Pearson matching we approximate these integrals.

@@

Sklar's theorem provides a useful characterization of multivariate distributions through copulas. Yet the choice of copula-based simulation algorithm affects which joint distributions may be simulated. Even in low-dimensional spaces (e.g., $d=3$), there exist valid multivariate distributions with *feasible* Pearson correlation matrices that NORTA cannot match exactly [@LH75]. This occurs when the bivariate transformations are applied to find the input correlation matrix, yet when combined the resultant matrix is indefinite. These situations do occur, even using exact analytic calculations. Such problematic target correlation matrices are termed *NORTA defective*.

@@

There is hope, however, as @GH02 also showed that replacing an indefinite input correlation matrix with a close proxy will give approximate matching to the target --- with adequate performance for moderate $d$. This provides evidence that our nearest positive definite (PD) augmented approach will maintain reasonable accuracy if our input matching scheme returns an indefinite matrix, at least for the rank-based matching scheme described above.

## Random vector generator {#rand-vec-gen}

We now describe our algorithm to generate random vectors, which mirrors the classical NORTA algorithm above with some modifications for rank-based dependency matching:

@@

1. Mapping step
    * (i) Convert the target Spearman correlation matrix $R_S$ to the corresponding MVN Pearson correlation $R_X$. Alternatively,
    * (i') Convert the target Kendall $\tau$ matrix $R_K$ to the corresponding MVN Pearson correlation $R_X$. Alternatively,
    * (i'') Convert the target Pearson correlation matrix to $R_P$ to the corresponding, approximate MVN Pearson correlation $R_X$.
2. Check admissibility and, if needed, compute the nearest correlation matrix.
    * (i) Check that $R_X$ is a correlation matrix, a positive definite matrix with 1's along the diagonal.
    * (ii) If $R_X$ is a correlation matrix, the input matrix is *admissible* in this scheme. Otherwise
    * (ii') Replace $R_X$ with the nearest correlation matrix $\tilde{R}_X$, in the Frobenius norm.
3. Gaussian copula
    * (i) Generate ${\bf X}=(X_1, \ldots, X_d) \sim N_d({\bf 0}, R_X)$.
    * (ii) Transform ${\bf X}$ to ${\bf U} = (U_1, \ldots,  U_d)$ viz. $U_i=\Phi(X_i)$, $i=1, \ldots, d$.
    * (iii) Return ${\bf Y}  = (Y_1, \ldots,  Y_d)$, where $Y_i=F_i^{-1}(U_i)$, $i=1, \ldots, d$.

*Step 1*

The first two descriptions of the *Mapping step* employ the closed-form relationships between $\rho_S$ and $\tau$ with $\rho_P$ for bivariate normal random variables via Equations \@ref(eq:convertKendall) and \@ref(eq:convertSpearman), respectively (implemented as `cor_covert`). Initializing our algorithm to match the nonparametric correlations by computing these equations for all pairs is computationally trivial.

For the computationally expensive process to match Pearson correlations, we approximate Equation \@ref(eq:pearsonIntegralRelation) for all pairs of margins. To this end, we implement the approximation scheme introduced by @xiao2019matching. Briefly, the many double integrals of the form in \@ref(eq:pearsonIntegralRelation) are approximated by weighted sums of Hermite polynomials. Matching coefficients for pairs of continuous distributions is made tractable by this method, but for discrete distributions (especially discrete distributions with large support sets or infinite support), the problem is not so simple or efficient.

Our solution is to approximate discrete distributions by a continuous distribution. The question then becomes that of which distribution to use. We found that Generalized S-Distributions [@muino2006gs] solve this problem by approximating a wide range of unimodal distributions, both continuous and discrete. Using the GS-Distribution approximation of discrete distributions in Pearson matching scheme above yields favorable results.

*Step 2*

Once $R_X$ has been determined, we check admissibility of the adjusted correlation via the steps described above. If $R_X$ is not a valid correlation matrix, then we compute the nearest correlation matrix. Finding the nearest correlation matrix is a common statistical computing problem. The defacto default function in `R` is `Matrix::nearPD`, an alternating projection algorithm due to @higham2002computing. As implemented the function fails to scale to HD. Instead, we provide the quadratically convergent algorithm based on the theory of strongly semi-smooth matrix functions (@qi2006quadratically). The nearest correlation matrix problem can be written down as the following convex optimization problem:

\begin{align*}
    \mathrm{min}\quad & \frac{1}{2} \Vert G - X \Vert^2 \\
    \mathrm{s.t.}\quad & X_{ii} = 1, \quad i = 1, \ldots , n, \\
    & X \in S_{+}^{n}
\end{align*}

For nonparametric correlation measures, our algorithm allows the generation of high-dimensional multivariate data with arbitrary marginal distributions with a broad class of admissible Spearman correlation matrices and Kendall $\tau$ matrices. The admissible classes consist of the matrices that map to a Pearson correlation matrix for a MVN. In particular, if we let $X$ be MVN with $d$ components and set

\begin{align*}
\Omega_P &= \{ R_P : R_P \textrm{ is a correlation matrix for } X \} \\
\Omega_K &= \{ R_K : R_K \textrm{ is a correlation matrix for } X \} \\
\Omega_S &= \{ R_S : R_S \textrm{ is a correlation matrix for } X \} \\
\end{align*}

then there are 1-1 mappings between these sets. We conjecture informally that the sets of admissible $R_S$ and $R_K$ are not highly restrictive. In particular, $R_P$ is approximately $R_S$ for a MVN, suggesting that the admissible set $\Omega_S$ should be flexible as $R_P$ can be any PD matrix with 1's along the diagonal. We provide methods to check whether a target $R_S$ is an element of the *admissible set* $\Omega_S$ (and similarly for $R_K$).

There is an increasing probability of encountering a non-admissible correlation matrix as dimension increases. In our experience, the mapping step for large $d$ almost always produces a $R_X$ that is not a correlation matrix. In Section \@ref(package) we provide a basic description of how to quantify and control the approximation error. Further, the RNA-seq example in Section \@ref(examples) provides an illustration of this in practice.

*Step 3*

The final step implements a NORTA-inspired, Gaussian copula approach to produce the desired margins. Steps 1 and 2 determine the MVN Pearson correlation values that will eventually match the target correlation. Then all that is required is a fast MVN simulator, a standard normal CDF, and well-defined quantile functions for marginals. The MVN is transformed to a copula (distribution with standard uniform margins) by applying the normal CDF $\phi(\cdot)$. The quantile functions $F^{-1}$ are applied across the margins to return the desired random vector ${\bf Y}$.

>>>>

This section describes our methods involved in simulating a random vector $\bf Y$ with $Y_i$ components for $i=1,\ldots,d$. Each $Y_i$ has a specified marginal CDF $F_i$ and its inverse $F^{-1}_i$. To characterize dependency, every pair $(Y_i, Y_j)$ has a given Pearson correlation $\rho_P$, Spearman correlation $\rho_S$, and/or Kendall's $\tau$. The method can be described as a *high-performance Gaussian copula* (Equation \@ref(eq:gauss)) providing a HD NORTA-inspired algorithm.

@@

1. Simulate a random vector $\bf Z$ with $d$ independent and identically distributed (iid) standard normal components.
2. Determine the input matrix $\Sigma_{\bf Z}$ that corresponds with the specified output $\Sigma_{\bf Y}$.
3. Produce a Cholesky factor $M$ of $\Sigma_{\bf Z}$ such that $M M^{\prime}=\Sigma_{\bf Z}$.
4. Set $X$ by $X \gets MZ$.
5. $\text{Return} \; Y \; \text{where} \; Y_i \gets F_i^{-1}[\Phi(X_i)], \; i=1,...,d$.

@@

With modern parallel computing, steps 1, 3, 4, 5 are readily implemented as high-performance, multi-core and/or graphical-processing-unit (GPU) accelerated algorithms --- providing fast scalability to HD.

Matching specified Pearson correlation coefficients exactly (step 2 above), however, is computationally costly. In general, there is no convenient correspondence between the components of the input $\Sigma_{\bf Z}$ and target $\Sigma_{\bf Y}$. Matching the correlations involves evaluating or approximating $\binom{d}{2}$ integrals of the form

\begin{equation}
    \mathrm{E}\left[Y_i Y_j\right] = \int_{-\infty}^{\infty} \int_{-\infty}^{\infty} F_i^{-1}\left[\Phi(z_i)\right] F_j^{-1}\left[\Phi(z_j)\right] \phi(z_i, z_j, \rho_z) dz_i dz_j,
    \label{eq:pearsonIntegralRelation}
\end{equation}

where $\phi(\cdot)$ is the joint probability distribution function of two correlated standard normal variables. For HD simulation, exact evaluation may become too costly to enable practical simulation studies. Our remedy to enable HD Pearson matching in `bigsimr` is Hermite polynomial approximation of these $\choose{d}{2}$ double integrals, as suggested by @XZ19.

@@

Sklar's theorem provides a useful characterization of multivariate distributions through copulas. Yet the choice of copula-based simulation algorithm affects which joint distributions may be simulated. Even in LD spaces (e.g., $d=3$), there exist valid multivariate distributions with *feasible* Pearson correlation matrices that NORTA cannot match exactly [@LH75]. This occurs when the bivariate transformations are applied to find the input correlation matrix, yet when combined the resultant matrix is indefinite. These situations do occur, even using exact analytic calculations. Such problematic target correlation matrices are termed *NORTA defective*.

@@

There is hope, however, as @GH02 also showed that replacing an indefinite input correlation matrix with a close proxy will give approximate matching to the target --- with adequate performance for moderate $d$. This provides guidance that our nearest positive definite (PD) augmented approach has promise provide adequate accuracy if our input matching scheme returns an indefinite matrix.

## bigsimr rvec: Random vector generator {#rand-vec-gen}

We now describe our algorithm to generate random vectors, which extends NORTA to HD via nearest correlation matrix replacement and provides rank-based dependency matching:

@@

1. Mapping step, either
    * Convert the target Spearman correlation matrix $R_S$ to the corresponding MVN Pearson correlation $R_X$. Alternatively,
    * Convert the target Kendall $\tau$ matrix $R_K$ to the corresponding MVN Pearson correlation $R_X$. Alternatively,
    * Convert the target Pearson correlation matrix to $R_P$ to the corresponding, approximate MVN Pearson correlation $R_X$.
2. Check admissibility and, if needed, compute the nearest correlation matrix.
    * Check that $R_X$ is a correlation matrix, a positive definite matrix with 1's along the diagonal.
    * If $R_X$ is a correlation matrix, the input matrix is *admissible* in this scheme. Otherwise,
    * Replace $R_X$ with the nearest correlation matrix $\tilde{R}_X$, in the Frobenius norm.
3. Gaussian copula
    * Generate ${\bf X}=(X_1, \ldots, X_d) \sim N_d({\bf 0}, R_X)$.
    * Transform ${\bf X}$ to ${\bf U} = (U_1, \ldots,  U_d)$ viz. $U_i=\Phi(X_i)$, $i=1, \ldots, d$.
    * Return ${\bf Y}  = (Y_1, \ldots,  Y_d)$, where $Y_i=F_i^{-1}(U_i)$, $i=1, \ldots, d$.

*Step 1: Mapping step.* We first employ the closed-form relationships between $\rho_S$ and $\tau$ with $\rho_P$ for bivariate normal random variables via Equations \@ref(eq:convertKendall) and \@ref(eq:convertSpearman), respectively (implemented as `cor_covert()`). Initializing our algorithm to match the nonparametric correlations by computing these equations for all pairs is computationally trivial.

For the computationally expensive process to match Pearson correlations, we approximate Equation \@ref(eq:pearsonIntegralRelation) for all pairs of margins. To this end, we implement the approximation scheme introduced by @xiao2019matching. Briefly, the many double integrals of the form in \@ref(eq:pearsonIntegralRelation) are approximated by weighted sums of Hermite polynomials. Matching coefficients for pairs of continuous distributions is made tractable by this method, but for discrete distributions (especially discrete distributions with large support sets or infinite support), the approximation is computationally expensive. To remedy this, we further approximate discrete distributions by a continuous distribution using Generalized S-Distributions [@muino2006gs].

*Step 2: Admissibility check and nearest correlation matrix computation.* Once $R_X$ has been determined, we check admissibility of the adjusted correlation via the steps described above. If $R_X$ is not a valid correlation matrix, then we compute the nearest correlation matrix. Finding the nearest correlation matrix is a common statistical computing problem. The defacto function in `R` is `Matrix::nearPD`, an alternating projection algorithm due to @higham2002computing. As implemented the function fails to scale to HD. Instead, we provide the quadratically-convergent algorithm based on the theory of strongly semi-smooth matrix functions (@qi2006quadratically). The nearest correlation matrix problem can be written down as the following convex optimization problem:  $\mathrm{min} \quad \frac{1}{2} \Vert G - X \Vert^2, \quad \mathrm{s.t.} \quad X_{ii} = 1, \quad i = 1, \ldots , n, \quad X \in S_{+}^{n}$.

For nonparametric correlation measures, our algorithm allows the generation of HD multivariate data with arbitrary marginal distributions with a broad class of admissible Spearman correlation matrices and Kendall $\tau$ matrices. The admissible classes consist of the matrices that map to a Pearson correlation matrix for a MVN. In particular, if we let $X$ be MVN with $d$ components and denote $\Omega_P = \{ R_P : R_P \textrm{ is a correlation matrix for } X \}, \quad \Omega_K = \{ R_K : R_K \textrm{ is a correlation matrix for } X \}, \quad \Omega_S = \{ R_S : R_S \textrm{ is a correlation matrix for } X \}$. There are 1-1 mappings between these sets. We conjecture that the sets of admissible $R_S$ and $R_K$ are not highly restrictive. In particular, $R_P$ is approximately $R_S$ for a MVN, suggesting that the admissible set $\Omega_S$ should be flexible as $R_P$ can be any PD matrix with 1's along the diagonal. We provide methods to check whether a target $R_S$ is an element of the *admissible set* $\Omega_S$.

There is an increasing probability of encountering a inadmissible correlation matrix as dimension increases. In our experience, the mapping step for large $d$ almost always produces a $R_X$ that is not a correlation matrix. In Section \@ref(package) we provide a basic description of how to quantify and control the approximation error. Further, the RNA-seq example in Section \@ref(examples) provides an illustration of this in practice.

*Step 3: Gaussian copula.* The final step implements a NORTA-inspired, Gaussian copula approach to produce the desired margins. Steps 1 and 2 determine the MVN Pearson correlation values that will eventually match the target correlation. Step 3 requires a fast MVN simulator, a standard normal CDF, and well-defined quantile functions for marginals. The MVN is transformed to a copula (distribution with standard uniform margins) by applying the normal CDF $\Phi(\cdot)$. Finally, the quantile functions $F^{-1}$ are applied across the margins to return the desired random vector ${\bf Y}$.

### 2021-05-25 060-conclusion-discussion.Rmd
Revise for JSS submission up to Chapter 4

We have introduced a general-purpose high-dimensional multivariate simulation algorithm and provide a user-friendly, high-performance `R` package [`bigsimr`]. The random vector generation method is inspired by NORTA [@Cario1997] and Gaussian copula-based approaches [@MB13, @BF17, @Xia17]. The major contributions of this work are methods and software for flexible, scalable simulation of HD multivariate probability distributions with broad potential data analytic applications for modern, big-data statistical computing. For example, one could simulate high-resolution time series data, such as those consistent with an auto-regressive moving average model exhibiting a specified Spearman structure. Or our methods could be used to simulate sparsely correlated data, as many HD methods assume, via specifying a *spiked correlation matrix*.

It is customary to compare new tools and algorithms directly to existing competing methods and software. In this study, however, we only employ our proposed methodology, as our previous work has shown that existing `R` tools are simply not designed to meet our high-dimensional goal (see @Li2019gpu for evaluations of the `R` `copula` package and others). For the bivariate simulations, existing packages such as [`nortaRA`](https://github.com/superdesolator/NORTARA) work well to match Pearson correlations exactly.

>>>>

We introduced a general-purpose high-dimensional multivariate simulation algorithm and provide a user-friendly, high-performance `R` package [`bigsimr`]. The random vector generation method is inspired by NORTA [@Cario1997] and Gaussian copula-based approaches [@MB13, @BF17, @Xia17]. The major contributions of this work are methods and software for flexible, scalable simulation of HD multivariate probability distributions with broad potential data analytic applications for modern, big-data statistical computing. For example, one could simulate high-resolution time series data, such as those consistent with an auto-regressive moving average model exhibiting a specified Spearman structure. Or our methods could be used to simulate sparsely correlated data, as many HD methods assume, via specifying a *spiked correlation matrix*.

<!-- 
It is customary to compare new tools and algorithms directly to existing competing methods and software. In this study, however, we only employ our proposed methodology, as our previous work has shown that existing `R` tools are simply not designed to meet our high-dimensional goal (see @Li2019gpu for evaluations of the `R` `copula` package and others). For the bivariate simulations, existing packages such as [`nortaRA`](https://github.com/superdesolator/NORTARA) work well to match Pearson correlations exactly.
-->

### 2021-05-25 article_bigsimr.Rmd
Revise for JSS submission up to Chapter 4

---
documentclass: jss
author:
  - name: Alfred G. Schissler
    affiliation: University of Nevada
    address: |
      | 1664 N Virginia St.
      | Reno, NV 89557
    email: \email{aschissler@unr.edu}
  - name: Edward J. Bedrick
    affiliation: University of Nevada
  - name: Alexander D. Knudson
    affiliation: University of Nevada \AND
  - name: Tomasz J. Kozubowski
    affiliation: University of Nevada
  - name: Tin Nguyen
    affiliation: University of Nevada
  - name: Anna K. Panorska
    affiliation: University of Nevada \AND
  - name: Juli Petereit
    affiliation: University of Nevada
  - name: Walter W. Piegorsch
    affiliation: University of Nevada
  - name: Duc Tran
    affiliation: University of Nevada
title:
  formatted: "Simulating High-Dimensional Multivariate Data using the \\pkg{Bigsimr} Package"
  plain:     "Simulating High-Dimensional Multivariate Data using the Bigsimr Package"
  short:     "\\pkg{Bigsimr}: Simulate High-Dimensional Data"
abstract: >
  It is critical to realistically simulate data when employing Monte Carlo techniques and evaluating statistical methodology. Measurements are often correlated and high dimensional in this era of big data, such as data obtained through high-throughput biomedical experiments. Due to computational complexity and a lack of user-friendly software available to simulate these massive multivariate constructions, researchers resort to simulation designs that posit independence or perform arbitrary data transformations. This article introduces the \\pkg{Bigsimr} R package. This high-level package provides a flexible and scalable procedure to simulate high-dimensional random vectors with arbitrary marginal distributions and known Pearson, Spearman, or Kendall correlation matrix. \\pkg{Bigsimr} contains additional high-performance features, including multi-core algorithms to simulate random vectors, estimate correlation, and compute the nearest correlation matrix. Monte Carlo studies quantify the accuracy and scalability of our approach, up to $d=10,000$. Finally, we demonstrate example applications enabled via \\pkg{Bigsimr} by applying these methods to a motivating dataset --- RNA-sequencing data obtained from breast cancer tumor samples.
keywords:
  formatted: [multivariate simulation, high-dimensional data, nonparametric correlation, Gaussian copula, RNA-sequencing data, breast cancer]
  plain:     [multivariate simulation, high-dimensional data, nonparametric correlation, Gaussian copula, RNA-sequencing data, breast cancer]
bibliography: [bigsimr.bib, packages.bib, alex.bib]
biblio-style: jss
preamble: >
  \usepackage{amsmath}
  \usepackage{amssymb}
  \usepackage{amsfonts}
  \usepackage{amsthm}
  \newtheorem{theorem}{Theorem}
  \newtheorem{lemma}[theorem]{Lemma}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{array}
  \usepackage{ragged2e}
  \usepackage{setspace}
output: rticles::jss_article
---

>>>>

---
documentclass: jss
author:
  - name: Alfred G. Schissler
    affiliation: University of Nevada
    address: |
      | 1664 N Virginia St.
      | Reno, NV 89557
    email: \email{aschissler@unr.edu}
  - name: Edward J. Bedrick
    affiliation: University of Arizona
  - name: Alexander D. Knudson
    affiliation: University of Nevada \AND
  - name: Tomasz J. Kozubowski
    affiliation: University of Nevada
  - name: Tin Nguyen
    affiliation: University of Nevada
  - name: Anna K. Panorska
    affiliation: University of Nevada \AND
  - name: Juli Petereit
    affiliation: University of Nevada
  - name: Walter W. Piegorsch
    affiliation: University of Arizona
  - name: Duc Tran
    affiliation: University of Nevada
title:
  formatted: "Simulating High-Dimensional Multivariate Data using the \\pkg{bigsimr} Package"
  plain:     "Simulating High-Dimensional Multivariate Data using the bigsimr Package"
  short:     "\\pkg{bigsimr}: Simulate High-Dimensional Data"
abstract: >
  It is critical to accurately simulate data when employing Monte Carlo techniques and evaluating statistical methodology. Measurements are often correlated and high dimensional in this era of big data, such as data obtained through high-throughput biomedical experiments. Due to computational complexity and a lack of user-friendly software available to simulate these massive multivariate constructions, researchers resort to simulation designs that posit independence or perform arbitrary data transformations. To meet this gap, we developed the bigsimr R package for high-dimensional random vector simulation with arbitrary marginal distributions and known Pearson, Spearman, or Kendall correlation matrix. bigsimr contains additional high-performance features, including multi-core algorithms to estimate correlation and compute the nearest correlation matrix. Monte Carlo studies quantify the accuracy and scalability of our approach, up to $d=10,000$. Finally, we describe example workflows and apply to a motivating high-dimensional data set --- RNA-sequencing data obtained from breast cancer tumor samples.
keywords:
  formatted: [multivariate simulation, high-dimensional data, nonparametric correlation, Gaussian copula, RNA-sequencing data, breast cancer]
  plain:     [multivariate simulation, high-dimensional data, nonparametric correlation, Gaussian copula, RNA-sequencing data, breast cancer]
bibliography: [bigsimr.bib, packages.bib, alex.bib]
biblio-style: jss
preamble: >
  \usepackage{amsmath}
  \usepackage{amssymb}
  \usepackage{amsfonts}
  \usepackage{amsthm}
  \newtheorem{theorem}{Theorem}
  \newtheorem{lemma}[theorem]{Lemma}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{array}
  \usepackage{ragged2e}
  \usepackage{setspace}
output: rticles::jss_article
---
